---
title: "Week 2 Classes"
author: "yob"
date: "5/14/2019"
output:
  html_document:
    df_print: paged
---

## Coding Standards in R

1- always use text files / text editor

2- indent your code

3- limit the width of your code to ~80 columns
select all, CMD + I
try not to do more than 2 nested for loops

4- limit the length of individual functions
one basic activity per function

## Markdown

markdown is a text-to-HTML conversion tool for web writers.

*This text will appear italicized!*

**Bold**

# main heading
## secondary heading
### tertiary heading

-first item in list
-second item in list
-third item in list

1. first item in list
2. second item in list
3. third item in list

links
[Download R](http://www.r-project.org/)

Advanced linking

I spend so much time reading [R bloggers][1] and [Simply statistics][2]!

[1]: http://www.r-bloggers.com/ "R bloggers"
[2]: http://simplystatistics.org. "Simply Statistics"


Newlines require a double space after a line

john grubers website for the official markdown documentation

github markdown guide

# R Markdown

a core tool in literate statistical programming

R markdown can be converted to standard markdown using the knitr package in R

markdown can be converted to HTML using markdown package in R

**slidify** package is easy fix for converting to slides



My First R Markdown File 
------------------------

This is my first R Markdown file.

Here, we're going to load some data.

```{r}
library(datasets)

data(airquality)

summary(airquality)
```

let's first make a parts plot of the data. 

```{r}

pairs(airquality)

```


here's a regression model of ozone on wind, solar R, and temperature

```{r}
library(stats)
fit <- lm(Ozone ~ Wind + Solar.R + Temp, data = airquality)
summary(fit)
```

# Literate Statistical Programming with knitr
## Part 1

literate programs are weaved to produce human-readable format, and tangled to produce machine-readable format

save data in non-proprietary formats

the code needs to run to be able to create the final document

## Part 2

what is knitr?

an R package available on CRAN

supports RMarkdown, LaTeX, and HTML as documentation languages

can export to PDF, HTML

[you need to know markdown language](http://goo.gl/MUt9i5)

what is knitr good for?

-manuals
-short/medium-length technical documents
-tutorials
-reports
-data preprocessing documents/summaries

what is knitr not good for?

-very long research articles
-complex time-consuming computations
-documents that require precise formatting

## Part 3

```{r}
set.seed(1)
x <- rnorm(100)
mean(x)
```

code chunks can have names, which is useful when we start making graphics

```{r firstchunk}
print("hey")
```

## Part 4

.Rmd -> .md -> .html

```{r simulation, echo = FALSE}
set.seed(1)
x <- rnorm(100)
mean(x)
```
you won't see the code, you will only see the output here.

```{r, results = "hide"}
print("the results of this code chunk will be hidden")

```


the current time is `r time` My favourite random number is `r rand`.

you can use r codes inside sentences as above.

incorporating graphics

```{r scatterplot, fig.height = 4, echo = TRUE}
x <- rnorm(100); y <- x + rnorm(100, sd = 0.5)
par(mar = c(5, 4, 1, 1), las = 1)
plot(x, y, main = "My Simulated Data")

```


it actually embeds the entire image inside the HTML file itself.

not most efficient, but useful. everything is inside one file


```{r fitmodel}
library(datasets)
data(airquality)
fit <- lm(Ozone ~ Wind + Temp + Solar.R, data = airquality)

```

here is a table of regression coefficients.

```{r showtable, results = "asis"}
library(xtable)
xt <- xtable(summary(fit))
print(xt, type = "html")

```

## setting global options

```{r setoptions, echo = FALSE}
opts_chunk$set(echo = FALSE, results = "hide")
opts_chunk$set(echo = TRUE, results = "asis")
```

output:
results : "asis" , "hide"
echo : TRUE, FALSE

figures
fig.height: numeric
fig.width: numeric

## caching computations

what is one chunk takes a long time to run?

the cache = TRUE option can be set on a chunk-by-chunk basis to store results of computation

it will load the results from the disk if nothing is changed.

if the data or the code changes, you need to re-run the cached code chunks

dependencies are not checked explicitly. be careful.

